Formalizing Stack Safety as a Security Property

### 2021-09-04 submission3/response.txt

Review #416A
Overall merit

B. Weak Accept
Reviewer expertise

X. Expert
Paper summary

This submission formalizes the general notion of stack safety (roughly speaking, every caller should be protected from its callees during program execution) using well-known notions of security (integrity, confidentiality and non interference), and defines several security policies for protecting the stack. The formalization is validated on existing (and less general) micro-policies for stack safety (depth isolation, lazy tagging and lazy clearing) using property-based random testing. This evaluation of former policies detected some security violations in former and less precise works.

First, this submission introduces the main required security policies, in terms of traces of memory, register states and stepwise stack-safety: stepwise integrity (an execution step does not modify the memory slots that are reserved for a caller at a given depth), confidentiality (a callee is insensitive to the context of its call), stepwise confidentiality, temporal stack safety, and well-bracketed control-flow (calls and returns should be correctly nested). Then, the submission describes the requirements on the machine model (presence of a program counter, used to distinguish between the different stack-frames, and a mechanism for recognizing the return from a call) and the threat model. The machine model is abstract enough to represent different architectures.

Next, the submission details in a progressive way the formalization of machine behaviors and stack safety using semantic rules. The first formalization of the execution stack is simple, with a single stack and no sharing between callers and callees. Then the model is enriched with function parameters passed on the stack (by value or by reference), with callee-saves registers and with coroutines. Moreover, the first model is implemented in Coq and this implementation reuses an existing RISCV specification. The QuickChick tool is used to test the stepwise property and validate an existing enforcement mechanism (called depth isolation) of the literature. Furthermore, as the previous policy is conservative and rather slow to be practical, more permissive variants of confidentiality and integrity properties are formalized, resulting in an observational property that is again validated by testing.
Strengths

    well-written submission, well-explained ideas

    general formalization that addresses a real challenge and that is validated by testing

Weaknesses

    lack of discussion about the limitations of the work

    medium delta with respect to previous works

Comments for author

This submission is well-written, with well-explained ideas. Examples of C and assembly programs are used throughout the submission to illustrate the various définitions. The formalization is provided as a supplementary material together with the test cases. The techniques and results of the submission are interesting and general enough to be reused in different settings.

I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.

Why did you choose this RISCV specification? Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)? Would it be easy to adapt your work to another ISA specification? How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model? More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation? Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

Could you comment on the Coq formalization that is provided as a supplementary material? Was it straightforward to write a Coq program from a C++ program? Moreover, I looked at the Coq development and found some admitted lemmas in it.
Other comments

The supplementary material is not mentioned in the submission. Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').

The first example programs (Fig. 1 to 4) are C programs. The programs in Fig. 5 and 6 are qualified as C++ programs, but they could be C programs and they are close to the first programs. Moreover, the C syntax for passing parameters by reference is surprising. I would have written f(&x)(line 934) and void f(int *a)(line 944).

    (83) when THAT an
    (83) actiVation
    (210) We CAN this form
    (310, 886,895,947) The return instruction is confusing as the return type of fis void.
    (357) when THE state
    (389) are ARE
    (568) two many closing parentheses (twice)
    (626) we need REQUIRE
    (626) followS
    (652) calls. That
    (796) I was expecting m' in (m,p',ctx').
    (1005) lack of space between the last two rules
    (1118) conjunctIon

Questions for the response period

See my former questions related to 1) the Coq formalization, 2) the choice of a RISCV model and 3) the limitations of the approach

====================================================================================================================================

Review #416B
Overall merit

C. Weak Reject
Reviewer expertise

Y. Knowledgeable
Paper summary

This paper presents a semantics for formalising safety properties of stack safety enforcement systems. The paper uses a map to say which parts of the stack can currently be accessed and which can't, similar to the semantics of separation logic. It then connects this to a notion of call and return, that can be configured to enable different machine's implementations of functions.

The paper uses a systematic testing approach with QuickChick to show that a hardware enforcement mechanism actually provides stack safety, and find bugs that are injected into the approach. It is great that you found the bug in lazy approach, however, it is fairly obvious type of security issue where uninitialised data leads to something escaping.

The paper discusses extension, where parameter passing can use the stack, and passing references to structures.

The paper is primarily a semantics paper, though that seems fairly straightforward. The more interesting aspect of the paper is the more elaborate examples in Section 6 and 7, and the more elaborate extensions in 8. But I had not followed the earlier aspects well enough to appreciate those.

Overall, I didn't find the paper particularly compelling in its current form. It seems complex for what it achieves earlier on, and doesn't have clear enough exposition for the later parts.
Strengths

    Checks properties of stack safety using capabilities
    Interesting topic

Weaknesses

    Exposition doesn't connect the different aspects of the paper well.

    Seems overly complex in comparison to other relational reasoning papers.

Comments for author

I found the paper did not connect the section well. Section 4 presents a lot of details for a framework, which is then used in Section 5. Seem very disconnected? For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section. How does it relate? Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them. Is this because they are different or just missing exposition.

Line 320: defition

Line 329: I was really interested in address taken at this point, and whether the model could handle it. I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier

Line 349: How does this related to control flow integrity/guard, and Intel's CET?

Line 448: "If the predicate always holds" never holds?

Line 510: "isRet" is a relation, but you are using as a function.

Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?

Line 642: wrok

====================================================================================================================================

Review #416C
Overall merit

B. Weak Accept
Reviewer expertise

Z. No familiarity
Strengths

The work presents a compelling model for stack security, framed in terms of existing notions of integrity and confidentiality.

Models are sufficiently abstract and general to be transportable to a variety of mechanisms and machines.

Notions of observable stack safety capture a more idealized notion that can reflect realistic violations, rather than "false positives" due to the limitations of particular mechanisms. This is my favorite contribution of this work.

Some evidence that the approach can be tailored to complex control mechanisms, by showing how it applies to a simple fixed coroutine model.
Weaknesses

Validation is performed using random testing rather than formal proof. However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
Comments for author

line 13: variant of our properties and /does so/ correctly, and

line 94: what about tail calls?

line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.

line 123: "we extend the generation ..." I'm not sure what that means.

line 210: We /call/ this form...

line 293: What about exceptions?

line 563: all possible calls that /can/ be reached

### 2021-09-05 submission3/response.txt

> Review #416A
> Overall merit

> B. Weak Accept
> Reviewer expertise

> X. Expert
> Paper summary

> This submission formalizes the general notion of stack safety (roughly speaking, every caller should be protected from its callees during program execution) using well-known notions of security (integrity, confidentiality and non interference), and defines several security policies for protecting the stack. The formalization is validated on existing (and less general) micro-policies for stack safety (depth isolation, lazy tagging and lazy clearing) using property-based random testing. This evaluation of former policies detected some security violations in former and less precise works.

> First, this submission introduces the main required security policies, in terms of traces of memory, register states and stepwise stack-safety: stepwise integrity (an execution step does not modify the memory slots that are reserved for a caller at a given depth), confidentiality (a callee is insensitive to the context of its call), stepwise confidentiality, temporal stack safety, and well-bracketed control-flow (calls and returns should be correctly nested). Then, the submission describes the requirements on the machine model (presence of a program counter, used to distinguish between the different stack-frames, and a mechanism for recognizing the return from a call) and the threat model. The machine model is abstract enough to represent different architectures.

> Next, the submission details in a progressive way the formalization of machine behaviors and stack safety using semantic rules. The first formalization of the execution stack is simple, with a single stack and no sharing between callers and callees. Then the model is enriched with function parameters passed on the stack (by value or by reference), with callee-saves registers and with coroutines. Moreover, the first model is implemented in Coq and this implementation reuses an existing RISCV specification. The QuickChick tool is used to test the stepwise property and validate an existing enforcement mechanism (called depth isolation) of the literature. Furthermore, as the previous policy is conservative and rather slow to be practical, more permissive variants of confidentiality and integrity properties are formalized, resulting in an observational property that is again validated by testing.
> Strengths

>    well-written submission, well-explained ideas

>    general formalization that addresses a real challenge and that is validated by testing

> Weaknesses

>    lack of discussion about the limitations of the work

>    medium delta with respect to previous works

> Comments for author

> This submission is well-written, with well-explained ideas. Examples of C and assembly programs are used throughout the submission to illustrate the various définitions. The formalization is provided as a supplementary material together with the test cases. The techniques and results of the submission are interesting and general enough to be reused in different settings.

> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.

*** Questions Begin ***

> Why did you choose this RISCV specification?
> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?

Because the enforcement mechanisms we test are existing micropolicies defined in terms of a reference monitor attached to a RISCV machine.
Note that the formalism itself is not dependent on the details of the architecture, only the testing framework to test that architecture is.
There is little in the tested micropolicies that depends on RISCV, either, but moving them to another architecture such as ARM would nevertheless
require translating the micropolicies rules to apply to that architecture.

> Would it be easy to adapt your work to another ISA specification?

Fairly easy. The formalism itself translates naturally to most ISA specifications. The largest amount of work
is in the testing framework - modifying the code generation system to match the new ISA.
We are currently working on extending the work to a model of a Cheri-style hardware capability machine, Cerise,
and have not encountered major obstacles.

> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?

Timing channels; anything else?

> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?

> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

I don't understand this question, any insight would be helpful.

> Could you comment on the Coq formalization that is provided as a supplementary material? Was it straightforward to write a Coq program from a C++ program? Moreover, I looked at the Coq development and found some admitted lemmas in it.

Writing Coq program from C++: I believe this one is for Rob
Admitted Lemmas:

> Other comments

> The supplementary material is not mentioned in the submission. Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').

The supplementary material should indeed be mentioned in the submission.
Explanation of definition 7.4:

> The first example programs (Fig. 1 to 4) are C programs. The programs in Fig. 5 and 6 are qualified as C++ programs, but they could be C programs and they are close to the first programs. Moreover, the C syntax for passing parameters by reference is surprising. I would have written f(&x)(line 934) and void f(int *a)(line 944).

>    (83) when THAT an
>    (83) actiVation
>    (210) We CAN this form
>    (310, 886,895,947) The return instruction is confusing as the return type of fis void.
>    (357) when THE state
>    (389) are ARE
>    (568) two many closing parentheses (twice)
>    (626) we need REQUIRE
>    (626) followS
>    (652) calls. That
>    (796) I was expecting m' in (m,p',ctx').
>    (1005) lack of space between the last two rules
>    (1118) conjunctIon

> Questions for the response period

> See my former questions related to 1) the Coq formalization, 2) the choice of a RISCV model and 3) the limitations of the approach

@@

====================================================================================================================================

> Review #416B
> Overall merit

> C. Weak Reject
> Reviewer expertise

> Y. Knowledgeable
> Paper summary

> This paper presents a semantics for formalising safety properties of stack safety enforcement systems. The paper uses a map to say which parts of the stack can currently be accessed and which can't, similar to the semantics of separation logic. It then connects this to a notion of call and return, that can be configured to enable different machine's implementations of functions.

> The paper uses a systematic testing approach with QuickChick to show that a hardware enforcement mechanism actually provides stack safety, and find bugs that are injected into the approach. It is great that you found the bug in lazy approach, however, it is fairly obvious type of security issue where uninitialised data leads to something escaping.

> The paper discusses extension, where parameter passing can use the stack, and passing references to structures.

> The paper is primarily a semantics paper, though that seems fairly straightforward. The more interesting aspect of the paper is the more elaborate examples in Section 6 and 7, and the more elaborate extensions in 8. But I had not followed the earlier aspects well enough to appreciate those.

> Overall, I didn't find the paper particularly compelling in its current form. It seems complex for what it achieves earlier on, and doesn't have clear enough exposition for the later parts.
> Strengths

>    Checks properties of stack safety using capabilities
>    Interesting topic

> Weaknesses

>    Exposition doesn't connect the different aspects of the paper well.

>    Seems overly complex in comparison to other relational reasoning papers.

> Comments for author

> I found the paper did not connect the section well. Section 4 presents a lot of details for a framework, which is then used in Section 5. Seem very disconnected? For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section. How does it relate? Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them. Is this because they are different or just missing exposition.

We should think if there is anything we can suggest here to improve organization, but it's unclear to me just what the problem was

> Line 320: defition

> Line 329: I was really interested in address taken at this point, and whether the model could handle it. I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier

> Line 349: How does this related to control flow integrity/guard, and Intel's CET?

> Line 448: "If the predicate always holds" never holds?

> Line 510: "isRet" is a relation, but you are using as a function.

> Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?

> Line 642: wrok

@@

====================================================================================================================================

> Review #416C
> Overall merit

> B. Weak Accept
> Reviewer expertise

> Z. No familiarity
> Strengths

> The work presents a compelling model for stack security, framed in terms of existing notions of integrity and confidentiality.

> Models are sufficiently abstract and general to be transportable to a variety of mechanisms and machines.

> Notions of observable stack safety capture a more idealized notion that can reflect realistic violations, rather than "false positives" due to the limitations of particular mechanisms. This is my favorite contribution of this work.

> Some evidence that the approach can be tailored to complex control mechanisms, by showing how it applies to a simple fixed coroutine model.
> Weaknesses

> Validation is performed using random testing rather than formal proof. However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
> Comments for author

> line 13: variant of our properties and /does so/ correctly, and

> line 94: what about tail calls?

Copy tail-call discussion from previous response

> line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.

> line 123: "we extend the generation ..." I'm not sure what that means.

> line 210: We /call/ this form...

> line 293: What about exceptions?

Copy exception discussion from previous response

> line 563: all possible calls that /can/ be reached

### 2021-09-06 submission3/response.txt

> Validation is performed using random testing rather than formal proof. However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
> Comments for author

> line 13: variant of our properties and /does so/ correctly, and

> line 94: what about tail calls?

Tail calls are a plausible extension of our framework, but not supported as written.
Since a tail call is both a call and a return in a single step, our normal handling of
returns will not work -- it would interpret the tail call as a call to and immediate return from
a new function activation. We would need to distinguish tail calls from normal calls and essentially
process the call and the return in the reverse order.

@@

> line 123: "we extend the generation ..." I'm not sure what that means.

> line 210: We /call/ this form...

> line 293: What about exceptions?

Exceptions are another plausible extension. They would involve interleaving the return
target stack with entries that are not return targets, but "catch targets." Reaching a catch target
would pop all return targets above it, much like returns that violate well-bracketedness already do.

### 2021-09-08 submission3/main.tex
Split StkTokens criteria and new criterion
Fixed a bunch of minor things

We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the depth of
the function activation that owns the frame and access to those locations is
then permitted only when that an actiation at that depth is currently executing, validates our
stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.
%
\ifaftersubmission
\bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}
\fi

@@

We can this form of confidentiality {\em stepwise confidentiality}. Informally,
for any state at the entry to a function, and for any variant over the values in the stack,
the primary state and the variant will step in lockstep until they
return. To step in lockstep
means that if both states step---that is, neither the primary nor the variant failstops---any
{\em system component} (such as memory cell, register, program counter) that changes
in the original will change in the same way in the variant.
When we later extend the model with a notion of observable output
(\cref{sec:lazy}), we will apply this condition only to those.

@@

Returns, on the other hand, need not be annotated: we can tell when a callee
has returned control to the caller when...we see control return to the
caller! We don't require the callee's cooperation at all.
For example, under a standard calling convention, we can see that a
return has occurred when state of the machine reaches a state where
the stack pointer has the same value as it did at the point of some earlier
call and the $\PCname$ points to the instruction following that call.  We
abstract away from the details of specific calling conventions by assuming
that we are given some way of telling when a given machine state is a
{\em return target} with respect to some set of pending calls.

@@

We adopt a strong threat model in which both caller and callee may execute
arbitrary machine code, subject to whatever well-formedness conditions the
enforcement mechanism demands. But hardware and timing attacks are out of scope, as are
are attacks on the heap---though we believe that the work of
\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
to protect both stack and heap.

@@

\begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f ~ \mpcstatename\)}
             {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \mpcstatename\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
    \judgment{}
             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
                {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \notfinished{\mpcstatename}{\MPCT'}\)}
  \end{minipage}
\end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate always holds, then
\(\PUT\) is the identity function.

@@

There remains the question of how to ensure that the sequences are
invoked at the right places from the perspective of the stack-safety
property. For a program to enjoy stack safety
(with respect to a particular call map), all we need require is that
the entry sequence immediately follow  any instruction marked as a call
in the call map. This is achieved by giving these instructions (another) special tag,
and it is easy to check statically that this has been done correctly.
If the code fails to initiate an exit
sequence at a point where the stack-safety property expects a return,
the micro-policy will incorrectly behave as if execution
is continuing in the callee, but since the callee never has stronger access
rights than the caller, this is harmless.

@@

There are several ways to evaluate whether an enforcement mechanism enforces
stack safety properties. Ideally such validation would be done through formal proof over
the semantics of the enforcement-augmented machine.
However, while there are no fundamental barriers to producing such a proof,
it would be considerable work to carry out for a full ISA like RISC-V and
complex enforcement mechanism like the Depth Isolation micro-policy.
We therefore choose to validate the micro-policy of the previous section by
systematically \emph{testing} that it satisfies our properties.
This focus on testing is better aligned with our immediate
goal of evaluating real enforcement mechanisms for real machine architectures.
Formal proof remains as possible future wrok.

@@

To use QuickChick, we build random test-case generators that produce
% \begin{enumerate}
% \item
(1) an initial RISC-V machine state, including most notably
  the program to be executed;
  % \item
(2)
  an initial policy state, tagging instructions corresponding to
  blessed call or return sequences appropriately, while marking
  all potential stack locations as $\tagNoDepth$; and
% \item
(3)
 a code map identifying the locations of calls
% \end{enumerate}
%
That is, our generators need to produce not only a low-level program
that exhibits interesting control-flow patterns, but also the policy
meta-information and code annotations that would normally be produced
by the compiler.

@@

\paragraphx{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of stack safety
is that of \citet{SkorstengaardSTKJFP}. Their
StkTokens system enforces stack safety on hardware equipped with
memory capabilities, similar to those used in the Cheri system~\citep{Woodruff+14,Chisnall+15}, but
with additional support for linear capabilities.
Stack safety is defined as the conjuncton of local state encapsulation (LSE) and
well-bracketed control flow (WBCF).
Informally, LSE means that the current activiation can only access its own frame,
and WBCF means that the topmost frame can only return to
the immediately adjacent frame below. Formally, these properties are defined
with reference to a capability machine with a built-in call stack and call and return
instructions; this machine is proven fully abstract with respect to a more
concrete capability machine, an approach termed ``fully abstract overlay semantics.''
The proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.

@@

\citet{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
candidate stack safety properties.
%% : they should be (1) intuitive; (2)
%% useful for reasoning; (3) reusable in secure compiler chains; (4)
%% arguably ``complete'' with respect to informal notions of stack
%% safety; and (5) potentially scalable to richer settings such as
%% dynamic code generation or multithreading.
%
(1) {\em Intuitiveness}. We contend that our property definitions are at least as
intuitive as theirs, while being phrased in interestingly different terms.
We use the language of formal security
properties to apply well-established concepts and formal techniques
of non-interference, and we define our properties in terms of
traces of memory and register states, rather than in terms of an
ad-hoc abstract machine.
Which approach is ``better'' is probably a matter of taste, but ours
has at least the merit of explicitly relating stack safety to existing
notions of security.
%
(2) {\em Usefulness} of the safety criterion
for carrying out formal proofs (about individual programs). We agree that this is an important consideration;
demonstrating it for our proposal remains future work, in which we
plan to apply our properties to justifying a low-level program logic.
%
(3) {\em Reusability} in secure compilation chains.
StkTokens
is specifically set up to address this,
while it is not a goal of
our work.
%
(4) {\em Comprehensiveness} in capturing programmers' informal
understanding of stack safety. We believe that our support
for dynamically changing notion of adversaries, as outlined in point (i) above,
captures a better definition of safety than LSE; moreover, our observational
properties capture an even richer and more extensional notion of stack safety
that can be applied to a broader range of enforcement mechanisms and code features.
%
(5) {\em Broad applicability} to richer settings. We have demonstrated that our approach extends to
coroutines, a first step towards multithreading. Dynamic code generation is
more challenging, but StkTokens does not yet handle this either.
%
Finally, we would propose one more criterion: having a generic definition
of stack safety that is directly reusable to evaluate multiple machines and
enforcement mechanisms. They have used their approach to give strong validation
to one machine/enforcement mechanism, a linear-capability machine,
via formal proof. We have used ours to obtain a weaker, testing-based validation of
two different mechanisms on a single machine, and will target a capability machine
next as future work.

>>>>

We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the depth of
the function activation that owns the frame and access to those locations is
then permitted only when an activation at that depth is currently executing, validates our
stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.
%
\ifaftersubmission
\bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}
\fi

@@

We call this form of confidentiality {\em stepwise confidentiality}. Informally,
for any state at the entry to a function, and for any variant over the values in the stack,
the primary state and the variant will step in lockstep until they
return. To step in lockstep
means that if both states step---that is, neither the primary nor the variant failstops---any
{\em system component} (such as memory cell, register, program counter) that changes
in the original will change in the same way in the variant.
When we later extend the model with a notion of observable output
(\cref{sec:lazy}), we will apply this condition only to those.

@@

Returns, on the other hand, need not be annotated: we can tell when a callee
has returned control to the caller when...we see control return to the
caller! We don't require the callee's cooperation at all.
For example, under a standard calling convention, we can see that a
return has occurred when the state of the machine reaches a state where
the stack pointer has the same value as it did at the point of some earlier
call and the $\PCname$ points to the instruction following that call.  We
abstract away from the details of specific calling conventions by assuming
that we are given some way of telling when a given machine state is a
{\em return target} with respect to some set of pending calls.

@@

We adopt a strong threat model in which both caller and callee may execute
arbitrary machine code, subject to whatever well-formedness conditions the
enforcement mechanism demands. But hardware and timing attacks are out of scope, as
are attacks on the heap---though we believe that the work of
\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
to protect both stack and heap.

@@

\begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f ~ \mpcstatename\)}
             {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \mpcstatename\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
    \judgment{}
             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
                {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \notfinished{\mpcstatename}{\MPCT'}\)}
  \end{minipage}
\end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds, then
\(\PUT\) is the identity function.

@@

There remains the question of how to ensure that the sequences are
invoked at the right places from the perspective of the stack-safety
property. For a program to enjoy stack safety
(with respect to a particular call map), all we require is that
the entry sequence immediately follows any instruction marked as a call
in the call map. This is achieved by giving these instructions (another) special tag,
and it is easy to check statically that this has been done correctly.
If the code fails to initiate an exit
sequence at a point where the stack-safety property expects a return,
the micro-policy will incorrectly behave as if execution
is continuing in the callee, but since the callee never has stronger access
rights than the caller, this is harmless.

@@

There are several ways to evaluate whether an enforcement mechanism enforces
stack safety properties. Ideally such validation would be done through formal proof over
the semantics of the enforcement-augmented machine.
However, while there are no fundamental barriers to producing such a proof,
it would be considerable work to carry out for a full ISA like RISC-V and
complex enforcement mechanism like the Depth Isolation micro-policy.
We therefore choose to validate the micro-policy of the previous section by
systematically \emph{testing} that it satisfies our properties.
This focus on testing is better aligned with our immediate
goal of evaluating real enforcement mechanisms for real machine architectures.
Formal proof remains as possible future work.

@@

To use QuickChick, we build random test-case generators that produce
% \begin{enumerate}
% \item
(1) an initial RISC-V machine state, including most notably
  the program to be executed;
  % \item
(2)
  an initial policy state, tagging instructions corresponding to
  blessed call or return sequences appropriately, while marking
  all potential stack locations as $\tagNoDepth$; and
% \item
(3)
 a code map identifying the locations of calls.
% \end{enumerate}

That is, our generators need to produce not only a low-level program
that exhibits interesting control-flow patterns, but also the policy
meta-information and code annotations that would normally be produced
by the compiler.

@@

\paragraphx{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of stack safety
is that of \citet{SkorstengaardSTKJFP}. Their
StkTokens system enforces stack safety on hardware equipped with
memory capabilities, similar to those used in the Cheri system~\citep{Woodruff+14,Chisnall+15}, but
with additional support for linear capabilities.
Stack safety is defined as the conjunction of local state encapsulation (LSE) and
well-bracketed control flow (WBCF).
Informally, LSE means that the current activiation can only access its own frame,
and WBCF means that the topmost frame can only return to
the immediately adjacent frame below. Formally, these properties are defined
with reference to a capability machine with a built-in call stack and call and return
instructions; this machine is proven fully abstract with respect to a more
concrete capability machine, an approach termed ``fully abstract overlay semantics.''
The proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.

@@

\citet{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
candidate stack safety properties.
%% : they should be (1) intuitive; (2)
%% useful for reasoning; (3) reusable in secure compiler chains; (4)
%% arguably ``complete'' with respect to informal notions of stack
%% safety; and (5) potentially scalable to richer settings such as
%% dynamic code generation or multithreading.
%
(1) {\em Intuitiveness}. We contend that our property definitions are at least as
intuitive as theirs, while being phrased in interestingly different terms.
We use the language of formal security
properties to apply well-established concepts and formal techniques
of non-interference, and we define our properties in terms of
traces of memory and register states, rather than in terms of an
ad-hoc abstract machine.
Which approach is ``better'' is probably a matter of taste, but ours
has at least the merit of explicitly relating stack safety to existing
notions of security.
%
(2) {\em Usefulness} of the safety criterion
for carrying out formal proofs (about individual programs). We agree that this is an important consideration;
demonstrating it for our proposal remains future work, in which we
plan to apply our properties to justifying a low-level program logic.
%
(3) {\em Reusability} in secure compilation chains.
StkTokens
is specifically set up to address this,
while it is not a goal of
our work.
%
(4) {\em Comprehensiveness} in capturing programmers' informal
understanding of stack safety. We believe that our support
for dynamically changing notion of adversaries, as outlined in point (i) above,
captures a better definition of safety than LSE; moreover, our observational
properties capture an even richer and more extensional notion of stack safety
that can be applied to a broader range of enforcement mechanisms and code features.
%
(5) {\em Broad applicability} to richer settings. We have demonstrated that our approach extends to
coroutines, a first step towards multithreading. Dynamic code generation is
more challenging, but StkTokens does not yet handle this either.

Finally, we would propose one more criterion: having a generic definition
of stack safety that is directly reusable to evaluate multiple machines and
enforcement mechanisms. They have used their approach to give strong validation
to one machine/enforcement mechanism, a linear-capability machine,
via formal proof. We have used ours to obtain a weaker, testing-based validation of
two different mechanisms on a single machine, and will target a capability machine
next as future work.
